---
title: "gh1"
author: "Min Kim"
date: "2025-02-03"
output: pdf_document
---

```{r}
library(here)
library(data.table)
library(readr)
```
## 4.1 
```{r, bring in the CLASSSUR dataset}
classsur <- read.table(here("Graded Homework", "classsur.txt"), header = TRUE, sep = "", fill = TRUE, stringsAsFactors = FALSE)
head(classsur)
summary(classsur)
```

### a.
Estimate the population mean of one of the measurement variables, such as age, grade point average (GPA), or study hours.

To estimate the population mean for a simple random sample of n accounts, we employ the sample average $$\bar{y}$$. 
$$
\bar{y} = \frac{\Sigma_{i=1}^n y_i}{n}
$$
```{r}
sum_age <- sum(classsur$Age) 
n <- nrow(classsur)
y_bar <- sum_age / n
print(y_bar)
```

### b.
Estimate a population proportion for one of the categorical variables, such as gender, class, or job status.

To estimate the population proportion, we employ $$\hat{p}$$.
```{r}
# Estimation population proportion of male samples / total number of samples
male <- sum(classsur$Gender == 1)
pop_prop_male <- male / n
print(pop_prop_male)
```

### c. 
Compare means on one variable for at least two different groups, such as men and
women.

We use the same variable, Gender, as part b) for part c). 
For a random sample from a population with independent random sample from a population with means $$\mu_y$$ and $$\mu_x$$, an unbiased estimate of $$mu_y - mu_x$$ is $$\bar{y}-\bar{x}$$. 

```{r}
mean_male <- mean(classsur$Gender == 1)
mean_female <- mean(classsur$Gender == 2)
diff_means <- mean_male - mean_female
print(diff_means)
```

### d.
Compare proportions on one categorical variable for at least two different groups
(i.e., class standing or location of permanent residence).

We compare the proportions of class standing. Class standing, provided in the Classsur dataset, is a categorical variable. Therefore, we treat classes 2.0, 3.0, 65.0, 4.0, 68.6, 1.0, 68.0, 61.0, 64.0, 67.0, 66.5, 108.0, and 5.0 as separate groups.  
```{r}
# identifying the unique groups in the Class standing variable
# unique(classsur$Class) 

classes_to_compare <- c(2.0, 4.0, 68.0)
class_prop <- prop.table(table(classsur$Class))
filtered_prop <- class_prop[names(class_prop) 
                            %in% classes_to_compare]
cat("proportions of classes 2.0, 4.0, and 68.0 each: ", filtered_prop)
```

## SRS 3
```{r, for SRS 3, we use the Turkey2011fv.csv dataset}

# read the Turkey2011fv.csv dataset, then check the structure of the csv file and a brief numerical summary
turkey2011 <- read.csv(here("Datasets", "Turkey2011fv.csv"))
str(turkey2011) 
summary(turkey2011)

# Population size N
N <- dim(turkey2011)[1]
cat("population size N: ", N)

# assign sample size n = 1500 as provided in problem instructions
n <- 1500

# filter data based on eftypes (1, 2, 3)
pop_1 <- subset(turkey2011, eftype == 1)$Nfraudmean
pop_2 <- subset(turkey2011, eftype == 2)$Nfraudmean
pop_3 <- subset(turkey2011, eftype == 3)$Nfraudmean
```
We estimate the difference of means of the number of eforensics-fraudulent votes (variable Nfraudmean) of two populations (eftype = 1, 2) based on srswor in the Turkey2011fv.csv file. From our computational calculation, we obtain the results as the following:
```{r}
# calculate means and standard errors
mean_1 <- mean(pop_1)
mean_2 <- mean(pop_2)
std_1 <- sd(pop_1)
std_2 <- sd(pop_2)

# compare the difference of means
diff_means <- mean_2 - mean_1

# standard error of difference
se_diff <- sqrt((std_1^2 / length(pop_1)) + (std_2^2 / length(pop_2)))

# construct a 95% CI
ci_lower <- diff_means - (1.96 * se_diff)
ci_upper <- diff_means + (1.96 * se_diff)

cat("Population 1 (eftype 1) Mean number of eforensics-fraudulent votes:", mean_1, "\n")
cat("Population 2 (eftype 2) Mean eforensics-fraudulent votes:", mean_2, "\n")
cat("Difference of Means:", diff_means, "\n")
cat("Standard Error of Differences of Means", se_diff, "\n")
cat("95% CI: [", ci_lower, ci_upper, "]", "\n")
```
Now, let us estimate the mean number of eforensics-fraudulent votes by type (eftype = 2, 3) using sample size n = 1500 and estimate the difference of means between the two types, evaluating the bound on the error of estimation as the following:
```{r}
# create subset populations for eftypes 2 and 3
pop_2 <- turkey2011$Nfraudmean[turkey2011$eftype == 2]
pop_3 <- turkey2011$Nfraudmean[turkey2011$eftype == 3]

# create random samples; get sample mean and stdev
set.seed(123)
sample_2 <- sample(pop_2, n, replace = FALSE)
sample_3 <- sample(pop_3, n, replace = FALSE)

mean_sample_2 <- mean(sample_2)
mean_sample_3 <- mean(sample_3)
sd2 <- sd(sample_2)
sd3 <- sd(sample_3)

cat("Mean for eftype == 2:", mean_sample_2, "\n")
cat("Mean for eftype == 3:", mean_sample_3, "\n")
cat("Standard Deviation for eftype = 2:", sd2, "\n")
cat("Standard Deviation for eftype = 3:", sd3, "\n")

# diff in sample means
diff_means <- mean_sample_3 - mean_sample_2

n2 <- length(sample_2[!is.na(sample_2)])
n3 <- length(sample_3[!is.na(sample_3)])

# standard error of difference
se_diff <- sqrt((sd2^2 / n2) + (sd3^2 / n3))

cat("Difference of Means:", diff_means, "\n")
cat("Standard Error of the Difference:", se_diff, "\n")

ci_lower <- diff_means - (1.96 * se_diff)
ci_upper <- diff_means + (1.96 * se_diff)

cat("Margin of Error:", 1.96 * se_diff)
cat("95% CI: [", ci_lower, ci_upper, "]", "\n")
```





